Automatic Metadata Generation for Fish Specimen Image Collections {\footnotesize \thanks{Research supported by NSF OAC \#1940233 and \#1940322.} }

### 2021-04-06 paper/paper_new.tex
Reworked Section III.

Since the digital collections may each contain tens of thousands of images, manually producing additional image quality metadata for each digitized specimen is both labor-and time intensive and prohibitively expensive. Methods for automatically computing metadata are therefore needed to fully exploit biological image repositories for scientific discovery.
%we are both extracting and calculating. Just to double check then, our definition/understanding of computing here includes both of these, or should we focus on extraction, with the idea that calculating is the extra novel thing added, the application of math-the groovy stuff you guys added.  
As a step towards improving metadata in specimen research image collections, members of Drexel University's Metadata Research Center are developing methods to automatically analyze fish images and
extract a set of data features that provide important metadata about the
digitized specimen.
The research is being conducted as part of the Biology Guided Neural Networks (BGNN) project, which aims to develop a novel class of artificial neural networks that exploit machine readable and predictive knowledge associated with specimen images, phylogenies and anatomy ontologies.
Using a combination of machine learning and image informatics 
techniques,
%tools and [okay to just say image informatics techniques instead of "image informatics tools and techniques" - thinking techniques implies tools.]
we can accurately determine general image quality and metadata, such as fish quantity and location within images, fish orientation, and
image scaling based on ruler identification and measurement. Image scaling
allows us to compute quantitative features about the fish specimens, such
as their length and area.  In order to test and validate our methods, they
were  applied to a set of \(7,244\)
images drawn from the Illinois Natural History Survey (INHS) Collection~\cite{INHS}.
Figure \ref{fig:INHS_examples} presents three typical images used in our
study.
%Dave/Joel, the section above, from "we can accurately ..." to the figures" to me is more results. I guess it's okay to give this in introduction. Just wanted to raise this point. I sometimes give a paraphrase too, but just good to see if it is a little too much or okay with you both. I think it's okay as I re-read it, but good to have someone check.
The following section of the paper provides contextual background for this work, followed by the research goals and objectives, and a review of our research methods. Next, the results are presented, followed by a discussion. The conclusion highlights key findings and identifies next steps.

@@

Our work stands apart from previous work in that we are
addressing a different need: lack of metadata is impeding scientific studies.
Our focus and motivation are different.
Effectively doing the computational pre-processing
that will enable further downstream analyses.

<--NOTE-NOTE-NOTE DAVE/JOEL/ALL - RE: above paragraph, I'm thinking we should we shift message a little bit to something like, e.g., "The research reviewed above demonstrates image features extraction; although researchers have not applied these approaches the rich and extensive collections of digitized specimens accessible in open repositories. Our research addresses this need by applying machine learning and informatics techniques to extract key metadata properties from the images, and, further computing higher-level metadata based on these initial results. Our goals and objectives are presented in the following section.

@@

Digitized specimen accessible in open access repositories provide a rich, extensive data source machine-learning and scientific discovery, however, these resources remain largely untapped due image quality and metadata limitations. The overall goal is addressing this need by combining machine-learning and image informatics techniques. The overall goal of our research is to offer a machine-driven [*novel/innovative?? blah...] approach to current manual metadata generation that is prohibitively costly in both terms of labor and time, and to further offer a [*novel] approach to computing higher-level metadata that supports scientific inquiry. Three key objectives are to:

GOAL 1) Test [*Explore, Demonstrate the us of] Facebook's AI Research's dectectron tool
Specific aims are to use dectectron to identify five object classes: fish, fish eyes, rulers, and the numbers 2 and 3 on rulers, as shown in Figure

GOAL 2) Pursue [*Investigate] image processing at the pixel level.
Pilot testing found the dectectron bounding boxes could be more tightly enclosed the detected objects. We will see if additional masking background as part of the fish, or the
bounding box may clip away part(s) of the fish with pixel analysis methods commonly
found in image informatics. Specific aim [*objectives] of our research are to:
%not sure we have to list all the details, but something like the below, with the 4 aspects of pixel analysis we tested/investigate/experimented with

1. Identify the appropriate threshold. We will compute optimal scaling between [*or is it Otsu thresholding] to identify an improved threshold value. [Or we will determine is we can computed as the halfway point between the Otsu threshold value and the mean of the background intensities.]

2. Consolidate the foreground

3. Adjusting the bounding box. [We will use computational methods for masking [calcuate a mask?] to determine if bounding box needs to be shrunk or expanded]
% Jane confused here, if this belongs here, or below

4. Developing a fallback approach for when flood fill operations behave unexpectedly the image is too washed out.

GOAL 3) Compute [*Test our ability to calculate] the following set of higher-level metadata properties: 1) contrast (intensities of the foreground and background, 2) centroid and eye center, and if eye is found, the fish 3) side 4) primary axis and clock value, finally the 5) scale and length.

@@

The automated metadata generation methods for our project were developed to work on a specific set of images, the INHS\ Fish Collection~\cite{INHS}.  Most of these images have been configured, produced and acquired with a standard procedure.  The images used for our study contain one fish placed on a bright, white background and contain an information tag and the same ruler.  See Figure \ref{fig:teaser} for two example images from the collection. While training and focusing our system on images with very similar compositions and visual properties may limit its immediate applicability, our efforts do demonstrate the potential that machine learning and image informatics techniques have for automatically generating metadata for biological specimen image collections.

@@

\begin{figure}[b]
  \centering
%  \includegraphics[width=.49\linewidth]{images/teaser1_crop}
  %\hspace{4mm}
  \includegraphics[width=.90\linewidth]{images/teaser2_crop}
  \caption{Initial object detection on a specimen image using Detectron2~\cite{wu2019detectron2}.}
  \label{fig:teaser}
\end{figure}

>>>>

Since the digital collections may each contain tens of thousands of images, manually producing additional image quality metadata for each digitized specimen is both labor-and time intensive and prohibitively expensive. Methods for automatically computing metadata are therefore needed to fully exploit biological image repositories for scientific discovery.
%we are both extracting and calculating. Just to double check then, our definition/understanding of computing here includes both of these, or should we focus on extraction, with the idea that calculating is the extra novel thing added, the application of math-the groovy stuff you guys added.  
As a step towards improving metadata in specimen research image collections, members of Drexel University's Metadata Research Center are developing methods to automatically analyze fish images and
extract a set of data features that provide important metadata about the
digitized specimen.
The research is being conducted as part of the Biology Guided Neural Networks (BGNN) project, which aims to develop a novel class of artificial neural networks that exploit machine readable and predictive knowledge associated with specimen images, phylogenies and anatomy ontologies.
Using a combination of machine learning and image informatics 
techniques,
%tools and [okay to just say image informatics techniques instead of "image informatics tools and techniques" - thinking techniques implies tools.]
we can accurately determine general image quality and metadata, such as fish quantity and location within images, fish orientation, and
image scaling based on ruler identification and measurement. Image scaling
allows us to compute quantitative features about the fish specimens, such
as their length and area.  In order to test and validate our methods, they
were  applied to a set of \(7,244\)
images drawn from the Illinois Natural History Survey (INHS) Collection~\cite{INHS}.
Figure \ref{fig:INHS_examples} presents three typical images used in our
study.
%Dave/Joel, the section above, from "we can accurately ..." to the figures" to me is more results. I guess it's okay to give this in introduction. Just wanted to raise this point. I sometimes give a paraphrase too, but just good to see if it is a little too much or okay with you both. I think it's okay as I re-read it, but good to have someone check.
The following section of the paper provides contextual background for this work, followed by the research goals and objectives, and a review of our research methods. Next, the results, along with discussion, are presented. The conclusion highlights key findings and identifies next steps.

@@

The research reviewed above involves image features extraction and
artificial intelligence; although the researchers have not applied these
approaches to the burgeoning collections of digitized specimens
accessible in open repositories. Our research addresses this need by
applying machine learning and image informatics techniques to extract key
metadata properties from fish specimen images and computing
higher-level metadata based on these initial results.
Our goals and objectives our presented below.

\begin{comment}
Our work stands apart from previous work in that we are
addressing a different need: lack of metadata is impeding scientific studies.
Our focus and motivation are different.
Effectively doing the computational pre-processing
that will enable further downstream analyses.

<--NOTE-NOTE-NOTE DAVE/JOEL/ALL - RE: above paragraph, I'm thinking we should we shift message a little bit to something like, e.g., "The research reviewed above demonstrates image features extraction; although researchers have not applied these approaches the rich and extensive collections of digitized specimens accessible in open repositories. Our research addresses this need by applying machine learning and informatics techniques to extract key metadata properties from the images, and, further computing higher-level metadata based on these initial results. Our goals and objectives are presented in the following section.
\end{comment}

@@

Digitized specimen accessible in open access repositories provide a rich,
extensive data source for machine-learning and scientific discovery.
These resources, however, remain largely untapped due to image quality
issues and metadata limitations.
The overall goal of our work addresses this need by combining
machine-learning and image informatics techniques to produce a computational
% . The overall goal of our research is to offer a machine-driven [*novel/innovative?? blah...]
alternative to the current manual metadata generation process, one that is
prohibitively costly in both terms of labor and time. Additionally, our
methods collectively provide a novel approach to computing higher-level
metadata that will support further scientific inquiry.

% [WHAT'S THE DIFFERENCE BETWEEN A GOAL AND AN OBJECTIVE?]

Four key objectives are to:

1) Explore use of Facebook AI Research's \verb|detectron| tool.  Specific
aims are to use \verb|detectron| to identify study-specific objects.
% classes: fish, fish eyes, rulers, and the numbers 2 and 3 on rulers, as shown in
% Figure \ref{fig:teaser}.

2) Investigate image processing at the pixel level.
Pilot testing found the detectron under-segmented the detected objects
with tightly enclosing bounding boxes. We will determine if
pixel analysis methods commonly found in image informatics may produce
more accurate bounding boxes and object masks.
% background as part of the fish, or the
% bounding box may clip away part(s) of the fish with
The specific aims of this objective are to:
%not sure we have to list all the details, but something like the below, with the 4 aspects of pixel analysis we tested/investigate/experimented with

\renewcommand{\labelenumi}{\alph{enumi}.}
\begin{enumerate}
\item
Identify the appropriate threshold value for more accurate mask.
% We will compute optimal scaling between [*or is it Otsu thresholding] to identify an improved threshold value. [Or we will determine is we can computed as the halfway point between the Otsu threshold value and the mean of the background intensities.]

\item
Remove noise to produce a single, solid mask.

\item
Compute more accurate bounding box from updated mask.
% Adjusting the bounding box. [We will use computational methods for masking [calcuate a mask?] to determine if bounding box needs to be shrunk or expanded]
% Jane confused here, if this belongs here, or below

\item
Automatically determine when modified methods fail and detectron values
should be used as is.
% Developing a fallback approach for when flood fill operations behave unexpectedly the image is too washed out.
\end{enumerate}

\begin{figure*}[t]
  \centering
 \includegraphics[width=.45\linewidth]{images/teaser1_crop}
  \hspace{5mm}
  \includegraphics[width=.45\linewidth]{images/teaser2_crop}
  \caption{Initial object detection on two specimen images using Detectron2~\cite{wu2019detectron2}.}
  \label{fig:teaser}
\end{figure*}

3) Compute a number of high-level metadata properties from the detected objects and image quality metrics.
% [*Test our ability to calculate] the following set of
% : 1) contrast (intensities of the foreground and background, 2) centroid and eye center, and if eye is found, the fish 3) side 4) primary axis and clock value, finally the 5) scale and length.

4) Compare computed metadata properties with manually generated properties
when possible to assess the accuracy and effectiveness of automated 
methods.

@@

The automated metadata generation methods for our project were developed to work on a specific set of images, the INHS\ Fish Collection~\cite{INHS}.  Most of these images have been configured, produced and acquired with a standard procedure.  The images used for our study contain one fish placed on a bright, white background and contain an information tag and the same ruler.  See Figure \ref{fig:INHS_examples} for three example images from the collection. While training and focusing our system on images with very similar compositions and visual properties may limit its immediate applicability, our efforts do demonstrate the potential that machine learning and image informatics techniques have for automatically generating metadata for biological specimen image collections.

@@
